---
params:
  printcode: false
  date: !r Sys.Date()
  year: 2022
  qtr: 2
  lob: 'OL Occ'
  model_path: './'
  max_ay: 2022
  loss_data_path: './'
  mcmc_func_path: './'
  plotting_script_path: './'
  documentation_script_path: './'
  first_ay: 2008
  n_ay: 15
  model: 'model.stan'
  n_chains: 6
  n_warmup: 1500
  n_iterations: 6000
  sel_lr: c(0.779, 0.263, 0.307, 0.326, 0.274, 0.350, 0.369, 0.363, 0.474, 0.479, 0.436, 0.513, 0.508, 0.695, 0.704)
  title: 'temp-title'
output:
  html_document:
    df_print: paged
title: "`r params$title`"
    
  
---

```{r Selected LR, message=F, echo=F, warning=F, }
sel_lr <- params$sel_lr
```


# Documentation for CSU OL Occurrence Model
### Data thru 2022Q2 

```{r Load Packages, echo=F, warning=F, message=F, cache=T}
### PACKAGES

library(tidyverse,quietly=TRUE)
library(stringr,quietly=TRUE)
library(bayesplot,quietly=TRUE)
library(ggplot2,quietly=TRUE)
library(rstan,quietly=TRUE)
library(loo,quietly=TRUE)
library(knitr,quietly=TRUE)

options(mc.cores=parallel::detectCores())
rstan_options(auto_write=TRUE)

params=list(year=year
                                    , qtr=quarter
                                    , lob="OL Occ"
                                    , model_path=model_path
                                    , max_ay=max_ay
                                    , loss_data_path=loss_data_path
                                    # , mcmc_func_path=mcmc_func_path
                                    # , plotting_script_path=plotting_script_path
                                    , documentation_script_path=documentation_script_path
                                    , n_ay=year - first_ay + 1
                                    , model=model                        ## stan model file
                                    , n_chains=n_chains                  ## number of independent Markov chains
                                    , n_warmup=n_warmup                  ## number of iterations at the start of each chain to discard
                                    , n_iterations=n_iterations           ## total number of iterations INCLUDING discarded warmup iterations
                                    # , title=rmd_title
                                    # , sel_lr=sel_lr
                                    )

filestem <- 'O:/PARM/Corporate Actuarial/Reserving/Ad Hoc Projects/202211 MCMC Modelling/CSU Analysis - 2022Q2 Data/'

# source('O:/PARM/Corporate Actuarial/Reserving/Ad Hoc Projects/202211 MCMC Modelling/CSU Analysis - 2022Q2 Data/scripts/mcmc_func.R')
# source('O:/PARM/Corporate Actuarial/Reserving/Ad Hoc Projects/202211 MCMC Modelling/CSU Analysis - 2022Q2 Data/scripts/plotting.R')
# source('O:/PARM/Corporate Actuarial/Reserving/Ad Hoc Projects/202211 MCMC Modelling/CSU Analysis - 2022Q2 Data/scripts/documentation.R')

source(r"(O:\PARM\Corporate Actuarial\Reserving\Scripts\R\mcmc_documentation\mcmc_documentation.R)")

```


```{r Load User Inputs, echo=F, warning=F, message=F, cache=T}
### USER INPUTS

## directory with the stan scripts
model.path <- params$model_path

## loss data
loss.data.path <- loss_data_path
```

## 1. Prep data  

```{r Build Data Sets, message=F, echo=F, warning=F, cache=T}

all_lines_olep <- build_csu_all_lines_olep_pt1(loss.data.path, as.character(params$year), as.character(params$qtr))
all_lines_df <- build_all_lines_csu_data(loss.data.path, as.character(params$year), as.character(params$qtr))
  
col.ord <- c('lob', 'ay', 'dev_month', 'cum_olep', 'rpt_loss', 'paid_loss', 'case_resv', 'paid_dcce', 'rpt_counts', 'closed_counts', 'open_counts', 'ave_rpt_loss', 'ave_paid_loss', 'ave_case_os')

modeldf <- build_model_df(loss.data.path, as.character(params$year), as.character(params$qtr), params$lob) %>%
  add_missing_data(min_ay=params$first_ay, n_w=params$n_ay, n_d=max(.$d)) %>%
  update_loss_prem_at_least_one(params$year, params$qtr) %>%
  update_counts_at_least_one() %>%
  recalculate_frequency_and_average_loss() %>%
  filter_by_calendar_idx(params$year, params$qtr) %>%
  mutate(a=as.numeric(w), b=as.numeric(d)) %>%
  arrange(lob, b, a) %>%
  select(-c(a, b)) %>%
  filter(lob==params$lob)

cur_year <- build_cur_year_df(modeldf, params)

dm_ord <- modeldf$dev_month %>% unique() %>% sort() %>% as.character()
```

### 1.1 Reported Loss  

```{r Output Reported Loss Triangle, message=F, echo=F, warning=F, cache=T}
output_triangle(modeldf, "rpt_loss", dm_ord)
```

### 1.2 Paid Loss  

```{r Output Paid Loss Triangle, message=F, echo=F, warning=F, cache=T}
output_triangle(modeldf, "paid_loss", dm_ord)
```


### 1.3 Paid DCCE  
```{r Output Paid DCCE Triangle, message=F, echo=F, warning=F, cache=T}
output_triangle(modeldf, "paid_dcce", dm_ord)
```


### 1.4 Cumulative Losses  
```{r Output table of cumulative losses by AY, message=F, echo=F, warning=F, cache=T}
rpt <- output_cumulative(modeldf, "rpt_loss", returndf=T)
pd <- output_cumulative(modeldf, "paid_loss", returndf=T)
dcce <- output_cumulative(modeldf, "paid_dcce", returndf=T)

cumdf <- rpt %>% 
  left_join(pd, by=c('lob', 'ay', 'dev_month')) %>%
  left_join(dcce, by=c('lob', 'ay', 'dev_month'))

cumdf$ay <- cumdf$ay %>% as.factor()
cumdf$dev_month <- cumdf$dev_month %>% as.factor()
cumdf$lob <- cumdf$lob %>% str_remove(" ")

names(cumdf) <- c('LOB', 'AY', 'Dev. Month', 'Cum. Reported Loss', 'Cum. Paid Loss', 'Cum. Paid DCCE')

options(knitr.kable.NA='')
cumdf %>% arrange(LOB, AY) %>% kable(format.args = list(big.mark=",", digits=0, scientific=F))
```


### 1.5 Build `stan_dat` object
```{r Building stan_dat object, message=F, echo=F, warning=F, cache=T}
## build data object to pass to stan (paid loss)
stan_dat <- list(len_data=dim(modeldf)[1]
                 , n_w=max(modeldf$w, na.rm=TRUE)
                 , n_d=max(modeldf$d, na.rm=TRUE)
                 , w=modeldf$w
                 , d=modeldf$d
                 , cur_d=cur_year$d
                 , log_prem=modeldf$cum_olep %>% log()
                 , log_prem_ay=cur_year$cum_olep %>% log()
                 , log_paid_loss=(modeldf$paid_loss + modeldf$paid_dcce) %>% log()
                 , log_rpt_loss=(modeldf$rpt_loss + modeldf$paid_dcce) %>% log()
                 , log_paid_loss_ay=(cur_year$paid_loss + cur_year$paid_dcce) %>% log()
                 , log_rpt_loss_ay=(cur_year$rpt_loss + cur_year$paid_dcce) %>% log()
                 )
```

```{r Fitting stan model, message=F, echo=F, warning=F, cache=F}
## fit the model
# dir <- r"(O:\PARM\Corporate Actuarial\Reserving\Ad Hoc Projects\202211 MCMC Modelling\CSU Analysis - 2022Q2 Data\output\paid_reported_v2.rda)"

model <- stan(data = stan_dat 
             , file = r"(O:\PARM\Corporate Actuarial\Reserving\Scripts\stan\paid_reported_v2.stan)"
             # , file = = stringr::str_c(model.path, params$model)
             , iter = params$n_iterations
             , warmup = params$n_warmup
             , chains = params$n_chains
             )

out_dir <- r"(O:\PARM\Corporate Actuarial\Reserving\Ad Hoc Projects\202211 MCMC Modelling\CSU Analysis - 2022Q2 Data\output\paid_reported_v2.rda)"
save(model, file=out_dir)



# model  <- run_if_file_exists(file=dir
#                              , ftnA = stan(data = stan_dat 
#                                            , file = r"(O:\PARM\Corporate Actuarial\Reserving\Scripts\stan\paid_reported_v2.stan)"
#                                            # , file = = stringr::str_c(model.path, params$model)
#                                            , iter = params$n_iterations
#                                            , warmup = params$n_warmup
#                                            , chains = params$n_chains
#                                            )
#                               , ftnB=load(file=dir)
# )
```

```{r Saving stan model, message=F, echo=F, warning=F, cache=T}
out_dir <- r"(O:\PARM\Corporate Actuarial\Reserving\Ad Hoc Projects\202211 MCMC Modelling\CSU Analysis - 2022Q2 Data\output\paid_reported_v2.rda)"
save(model, file=out_dir)
```

## 2. Results

### 2.1 Estimated Ultimates

## 2. Stan Model

### 2.1 Model Parameters

```{r Output model parameters, echo=F, message=F, warning=F, cache=T}
# k <- params %>% attr('names')
# v <- c()
# for(p in k){
#   v <- v %>% append(l[[p]][1])
# }
# mparadf <- data.frame(Parameter=k, Value=v)
# 
# kable(mparadf)
```

### 2.2 Stan code  

```{r Get stan code, message=F, echo=F, warning=F, cache=T}
model_str <- model %>% get_stancode()
```

#### 2.2.2 `functions` block 

```{r Output data code block, message=F, echo=F, warning=F, cache=T}
output_code_chunk(model_str, 'functions') %>% cat()
```

#### 2.2.2 `data` block 

```{r Output data code block, message=F, echo=F, warning=F, cache=T}
output_code_chunk(model_str, r"(data\{)") %>% cat()
```

#### 2.2.3 `transformed data` block
```{r Output transformed data code block, message=F, echo=F, warning=F, cache=T}
output_code_chunk(model_str, 'transformed data') %>% cat()
```

#### 2.2.4 `parameters` block 

```{r Output parameters code block, message=F, echo=F, warning=F, cache=T}
output_code_chunk(model_str, 'parameters') %>% cat()
```


#### 2.2.5 `transformed parameters` block  

```{r Output transformed parameters code block, message=F, echo=F, warning=F, cache=T}
output_code_chunk(model_str, 'transformed parameters') %>% cat()
```


#### 2.2.6 `model` block  

Note:  see appendix A.1 for model specification. 

```{r Output model code block, message=F, echo=F, warning=F, cache=T}
data_start <- model_str %>% str_locate(r"(model\{)")
data_end <- str_sub(model_str, start=data_start[[1, 1]]) %>% str_locate(r"(\})")
str_sub(model_str, start=data_start[[1, 1]]) %>% str_sub(start=1, end=data_end[2]) %>% cat()
```

#### 2.2.6 `generated quantities` block

```{r Output generated quantities code block, message=F, echo=F, warning=F, cache=T}
output_code_chunk(model_str, 'generated quantities') %>% cat()
```

## 3. Results

### 3.1 Estimated Reported Loss Next Quarter

```{r Output 1-Quarter Reported Loss Projection, message=F, echo=F, warning=F, cache=T}
res_df <- model %>% 
  transform_to_df() %>%
  select(starts_with(c('est_cum_paid_loss', 'est_cum_rpt_loss','est_ult_loss'))) %>%
  as.data.frame()

res2_df <- data.frame(param=names(res_df))
res2_df$std <- res2_df$param %>% sapply(function(x){res_df[, x] %>% sd()})
res2_df$mean <- res2_df$param %>% sapply(function(x){res_df[, x] %>% mean()})

for(q in c(0.25, 0.5, 0.75, 0.995)){
  res2_df[, str_c('q', as.character(q * 100))] <- res2_df$param %>% sapply(function(x){res_df[, x] %>% quantile(q)})
}

res2_df$IQR <- res2_df$q75 - res2_df$q25



res2_df %>% filter(grepl('rpt', param)) %>% kable(format.args = list(big.mark=",", digits=0, scientific=F))



```

### 3.2 Estimated Paid Loss Next Quarter

```{r Output 1-Quarter Paid Loss Projection, message=F, echo=F, warning=F, cache=T}
res2_df %>% filter(grepl('paid', param)) %>% kable(format.args = list(big.mark=",", digits=0, scientific=F))
```


### 3.3 Estimated Ultimates

```{r Output Estimated Ultimates, message=F, echo=F, warning=F, cache=T}
res2_df %>% filter(grepl('ult', param)) %>% kable(format.args = list(big.mark=",", digits=0, scientific=F))
```

## 4. Diagnostics

```{r Create plotdf of samples, cache=T}
plotdf <- model %>%
  transform_to_df() %>%
  select(starts_with(c('alpha_loss', 'beta_paid_loss', 'beta_rpt_loss', 'est_cum_paid_loss', 'est_cum_rpt_loss'
                       , 'paid_loss_resid', 'rpt_loss_resid', 'est_ult_loss', 'rho', 'gamma', 'speedup'))) %>%
  as.data.frame()


```

```{r Output beta, alpha, est_ult, est_cum, residual quantities}

df_filename <- stringr::str_c('../output/', lob %>% stringr::str_to_lower() %>% stringr::str_replace(' ', '_'), '.fst')

plotdf %>% 
  select(starts_with(c('alpha_loss', 'beta_paid_loss', 'beta_rpt_loss', 'est_cum_paid_loss', 'est_cum_rpt_loss', 'paid_loss_resid', 'rpt_loss_resid', 'est_ult_loss'))) %>%
  mutate(lob=params$lob) %>%
  fst::write_fst(path=df_filename, compress=80)

```


### 4.1 Stan model diagnostics  

#### 4.1.1 MCMC Trace plot

```{r Output alpha trace plots}
model %>%
  bayesplot::mcmc_trace(pars=c((params$n_ay - 8):params$n_ay) %>% 
                          sapply(function(x){str_c('alpha_loss[', x, ']')})
                        )
```
```{r Output beta_rpt trace plots, message=F, echo=F, warning=F}
model %>%
  bayesplot::mcmc_trace(pars=c(1:9) %>% sapply(function(x){str_c('beta_rpt_loss[', x, ']')}))
```
```{r Output beta_paid trace plots, message=F, echo=F, warning=F}
model %>%
  bayesplot::mcmc_trace(pars=c(1:9) %>% sapply(function(x){str_c('beta_paid_loss[', x, ']')}))
```


```{r Output est_ult_loss trace plots}
model %>%
  bayesplot::mcmc_trace(pars=c((params$n_ay - 8):params$n_ay) %>% 
                          sapply(function(x){str_c('est_ult_loss[', x, ']')})
                        )

```


```{r Distributions of last 4 AYs, message=F, echo=F, warning=F}
model %>%
  bayesplot::mcmc_areas(pars=c((params$n_ay-3):params$n_ay) %>% sapply(function(x){str_c('est_ult_loss[', x, ']')})) + ggtitle("Modelled Distribution of Ultimate Loss, Last Four Accident Years")
```
#### Distribution of Rhat Estimates

$\hat{R}<1.05$ is considered "good". Lighter colors are better.

```{r Distribution of Rhat estimates, message=F, echo=F, warning=F}
# bayesplot::mcmc_rhat_hist(Rhat(model))
```

#### Distribution of Effective Sample Sizes  

$\frac{\text{effective sample size}}{\text{total sample size}} $

Lighter colors are better.

```{r Neff ratios, message=F, echo=F, warning=F}
# bayesplot::mcmc_neff_hist(neff_ratio(model))
```

#### Distributions of Auxiliary Parameters  

```{r Distribution of Auxiliary Parameters, message=F, echo=F, warning=F}
model %>% 
  bayesplot::mcmc_dens_overlay(pars=c('rho', 'gamma'))
```

```{r, message=F, echo=F, warning=F}
# point_estimate <- model %>% 
#   get_stanmodel() %>%
#   optimizing(stan_dat)
# 
# point_estimate <- data.frame(value=point_estimate$par)
# row.names(point_estimate) <- NULL
```

```{r}
# model %>% as.array() %>% ess_bulk()
```


```{r Boxplots of Ultimate Losses}
ult_loss_boxplot(plotdf, sel_lr, cur_year, divide_by = 1000000)
```

```{r Boxplots of Report Patterns}
rpt_beta_boxplot(plotdf)
```

```{r Boxplots of Payment Patterns}
paid_beta_boxplot(plotdf)
```








## A. Appendix  

### A.1 Model Specification  

Assume a triangle $\mathbf{P}$ of cumulative paid losses and $\mathbf{R}$ of cumulative reported losses. For an accident period $w$ and development period $d$, denote the cumulative paid loss by $P_{w,d}$ and the cumulative reported loss by $R_{w,d}$. For each accident period $w$, denote the estimate of ultimate loss by $\hat{U}_w$, and the exposure by $e_w$.

Let $b_{wd}^R = \log (R_{w,d} / \hat{U}_w)$ be the percent of ultimate from accident period $w$ that has been *reported* as of development period $d$, and $b_{wd}^P = \log (P_{w,d} / \hat{U}_w)$ be the same thing for paid loss. For each accident period let $a_w$ be the expected *a priori* ultimate loss. Note that $a_w$ does not depend on $R$ or $P$, and $a_w$ may or may not be the same as the final $\hat{U}_w$, since $a_w$ is an a priori estimated ultimate loss.

We model $b_{wd}^R$ as $$E\left[ b_{wd}^R \right] = \beta_d^R$$ and $b_{wd}^P$ as $$E\left[ b_{wd}^P \right] = \beta_d^P$$ for all $w$. Thus each development pattern (either paid or reported) is assumed to be the same on average for every accident period. 

We thus start with the simple model
$$
\begin{align}
\beta_d^R &\sim N(0, \sqrt{10})  \\
\beta_d^P &\sim N(0, \sqrt{10})  \\
\alpha_w &\sim N(\log (e_w) + \log(\text{ELR}), \sqrt{10})

\end{align}
$$